The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-04-08 src/ms.tex
Lim edits for Sections 3.2-3.4

As part of the process of developing \astropypkg into a long-term sustainable
product, and to improve transparency and accountability, the Project agreed to
write down and formalize our governance structure (partly supported by explicit
funding for this purpose; see Section~\ref{sec:project-funding}). At the 2019
Astropy Coordination Meeting, input was gathered from participants on what governance
structures existed in the associated Open Source Software communities, and what
would fit well with the needs of \astropy. This led into a ``retreat'' planned
for March 2020, but due to the COVID-19 pandemic, this became a series of
virtual meetings of the ``Astropy Governance Working Group.'' This group
drafted the APE 0 document \citep{ape0}, which was then eventually ratified and
implemented by the ``Astropy Governance Implementation Working Group'' in Fall
2021. While the process emphasized flexibility and the ability to adapt to
changing circumstances, it is expected that this is the framework Astropy's
governance will operate in for at least the medium-term future.

The APE 0 \citep{ape0} document lays out the principles of this governance
structure, so we refer the reader to that document for a more thorough
description, and here we only highlight some key elements. While many of these
principles were already de facto true or have been discussed organically (and
have been discussed in earlier papers in this series), the APE 0-based governance
aims to provide a single place where the community can agree as a starting
point. With this in mind, it highlights the developer and user community of the
Astropy Project as the ultimate sources of authority, as well as the core
principle of ``do-ocracy'': those who do work for the Project (be it coding,
training, or other harder-to-quantify contributions) gain more influence on the
outputs of the Project by virtue of their effort. However, APE 0 adds the
concept of ``voting members'': a self-governed part of that community who are
entrusted to elect the Coordination Committee (CoCo). While the CoCo has existed
from the inception of the project, APE 0 establishes a formal voting process for
CoCo members, and explicitly outlines the rights and responsibilities of the
CoCo. This role is mainly to facilitate consensus and act as the decision maker
when other mechanisms have failed, and includes powers that either
require central authority or secrets (e.g., passwords). However, APE 0 also
charges the Committee to devolve responsibilities and seek community input on
these items as often as possible.

The first Coordination Committee election under these rules took place in Fall
2021, electing a mix of prior and new coordination committee members,
and was contested in the sense of more candidates than available slots. This
suggests the process is already working to serve the long-term interests of the
committee to both spread the coordination effort, and to ensure it is not
dominated by the same people for as long as the Project continues. While other,
more fine-grained governance improvements are planned for the future, it is
clear the foundation is now in place.

@@

\begin{figure}[t!]
    \begin{centering}
      \includegraphics[width=0.9\textwidth]{Ncommitters_vs_frac_commits.pdf}
        \caption{
            The cumulative fraction of Git commits to the \astropypkg core
            package as a function of number of contributors (committers) broken
            into three equal-length time periods between the start of \astropy
            and the feature freeze date of v5.0.
            The time eras early, mid, and recent correspond to the approximate
            year ranges 2011--2015, 2015--2018, 2018--2021 (defined more
            precisely in Section~\ref{sec:project-contributors}).
            In all eras, the majority of commits to the \astropypkg core package
            have come from a small subset of the total number of contributors.
            For example, as annotated in the figure, 90\% of the Git commits
            have come from 18, 34, and 44 committers in the early, mid, and
            recent eras, respectively.
        }
        \label{fig:contributor-summary:Ncommitters}
    \end{centering}
\end{figure}

As described in the introduction above, the \astropy Project is broader than the
core \astropypkg package as it also represents a number of core infrastructure
packages, affiliated packages, educational initiatives, and user communities.
In total, well over 1500 people have contributed in some way to the Project,
whether, for example, through contributing code or documentation, opening issues
on \github, providing infrastructure support, writing tutorials, or helping to
coordinate workshops that relate to Astropy.
The vast majority of these contributors and participants have interacted as
volunteers choosing to engage with the open-source and open-development software
community.
However, while the total number of contributors is large, the number of
contributions and bulk of the maintenance of the \astropy Project components has
primarily been sustained by a small number of heavily-engaged or
explicitly-employed contributors \citep[see, e.g., ][]{astropy:2018}, as is the
case in many large, open-source software communities
\citep[e.g.,][]{numpy:nature}.

@@

Since its inception, the most active contributors to the \astropypkg package
have spanned a range of career stages and career paths.
For example, Figure~\ref{fig:contributor-summary:top} shows a breakdown of the
career stage/path at the midpoint of the three time eras defined in the previous
paragraph for all contributors with more than 50 commits to the \astropypkg
package within each era.
Over the $\sim$6 years between the ``early'' and ``recent'' eras, many
contributors moved between these categories (i.e., have graduated as students to
become research staff / faculty or software engineers), but also a number of
early contributors moved on to other positions and stopped contributing to the
core \astropypkg package.
Given this, it is encouraging that the numbers in each career stage/path have
stayed relatively constant over the history of \astropy (with some showing
growth), as new, active contributors have generally joined and stepped in to
fill top contributor roles within the project.
On the other hand, the total number of top contributors has not grown
substantially, reflecting one of our principal challenges: sustaining a pipeline
from users to contributors to maintainers of the package.
One other thing to note in this context is that in all eras, 50--60\% of the
software engineers who contributed to \astropypkg were (or are) employed by
STScI, whose leadership has consistently provided resources (financial and
personnel) towards sustaining \astropypkg.

@@

% Much of this text was adapted from Robel Geda's report from NSBP and
% SACNAS. Please include him on the author list.
\textbf{\astropy representation at national diversity conferences:}
Several members of the \astropy community attended virtual conferences of the
National Society of Black Physicists (NSBP) and SACNAS (focusing on Hispanic and
Native American scientists across all STEM fields). \astropy representatives
noted that one underlying topic came up multiple times. A major barrier for
persons who come from underrepresented communities is the lack of resources and
expertise that are necessary to train students. Writing PEP 8-compliant software
and understanding the Git and \github workflow is not part of the standard
Physics and Astrophysics curriculum. Additionally, the larger astronomical
community is still in the process of transitioning towards \python-based tools,
making it difficult for students not currently under advisement by someone with
extensive \python expertise to get involved with the open-source \python
community. Projects such as Learn Astropy (see Section~\ref{sec:learn}) could
have a profound impact by empowering underrepresented groups because it provides
a free, searchable, and accessible introduction to \python tools for
astrophysical research. Representatives also noted that offering training and
teaching materials to PIs of Research Experience for Undergraduate (REU)
programs throughout the United States could be helpful for encouraging advisors
to teach students how to use Astropy, create their own libraries, and use
version control on open-source platforms like \github. Such materials could also
be offered as a workshop at national diversity conferences themselves, as SACNAS
solicits special session proposals each year.

@@

With the goal of establishing a sustainable financial model for the Astropy
Project, the Coordination Committee, under the new governance charter (see
Section~\ref{sec:project-governance}), established a standing Finance Committee.
This committee oversees the planning and allocation of finances on behalf
of the Project. The Astropy Project accepts funds from institutions as well as
individuals through NumFOCUS.\footnote{NumFOCUS is a 501(c)(3) nonprofit that
supports and promotes world-class, innovative, open-source scientific computing.}
Previously, no direct financial support was available for the project development,
and NumFOCUS covered most of the incurred operational costs. However, transitioning
to long-term financial stability meant having an influx of funds in the form of
institutional support, and hence the Astropy Project applied for and successfully
acquired four grants over the last several years. These grants are administered by NumFOCUS on behalf of the
Project, providing a ``neutral'' space not tied to any particular astronomy
institution. The two major grants are briefly summarized below.

\begin{description}
    \item[Moore] The Gordon and Betty Moore Foundation awarded the Astropy
    project a $\sim$\$900k (US) grant in 2019 for the proposal ``Sustaining and
    Growing the Astropy Project.'' Over a three-year term, this grant supported
    sustenance of the Project by providing funds for (1) the development and
    maintenance of the \astropypkg and its infrastructure, provided to existing
    project members and targeted hires, (2) the transition of long-term users to
    contributors and maintainers through mentorship efforts, (3) the
    formalization of a governance structure (see
    Section~\ref{sec:project-governance}), and (4) the improvement of equity,
    diversity, and inclusion efforts in the project (see
    Section~\ref{sec:project-ide}). Some of this money supported travel for
    meetings, conferences, and workshops. This generous grant helped reduce
    reliance on volunteer-driven maintenance of the openly-developed package and
    paved the way to acquire funding from US federal agencies, representing a
    major milestone for the success of the project.

    \item[NASA] A successful proposal to the NASA ROSES-2020
    call\footnote{http://solicitation.nasaprs.com/ROSES2020}, section E.7
    (Support for Open Source Tools, Frameworks, and Libraries), granted the
    Astropy Project $\sim$\$600k (US) to support its work. Funded work for a
    term of three years includes (1) project-wide infrastructure maintenance and
    improvements, (2) targeted work on areas of the \astropy core and
    coordinated packages, (3) enhancements to the Learn
    Astropy ecosystem and educational materials (see Section~\ref{sec:learn}),
    and (4) support for the \astropy affiliated packages. These goals
    support the sustenance of the project by encouraging further
    development of the package ecosystem and community engagement.
\end{description}

In addition to the two grants described above, the project benefited from
a Gemini Observatory contract under its Science User Support Department,
which awarded funds for development work that supports both the Astropy and
DRAGONS projects. The project also received financial support from the Dunlap
Seed Funding program at the University of Toronto to develop educational
resources for scientific software packaging within the Learn Astropy
framework. All of these successful proposals show that the Astropy Project
is vital for the astronomical infrastructure and community at large, and
strengthen its promising and sustainable future.

@@

As defined in \paperii, Astropy affiliated packages are astronomy-related
\python packages that provide functionality that builds upon and extends the
\astropypkg core package that have requested to be included as part of the
Astropy Project community. These packages support the goals and vision of
Astropy of improving code re-use, interoperability, and embracing good coding
practices such as testing and thorough documentation.
All Astropy-coordinated and other affiliated packages are listed with detailed
information on the \href{{https://www.astropy.org/affiliated}}{\astropy
website}; A table summarizing key aspects of the affiliated packages is included
in the Appendix below (Table~\ref{tab:affiliated-registry}).

Since \paperii, there has been an expansion of affiliated
packages for gravitational astrophysics, including:
\texttt{PyCBC} for exploring gravitational wave signals, \texttt{lenstronomy} for
modeling strong gravitational lenses, \texttt{ligo.skymap} for visualizing
gravitational wave probability maps, and \texttt{EinsteinPy} for general
relativity and gravitational astronomy. There have also been several packages
added to the ecosystem related to HEALPix: \texttt{astropy-healpix}
for a BSD-licensed HEALPix implementation, and \texttt{mocpy} for Multi-Order
Coverage maps. \texttt{astroalign} has been introduced for astrometric registration,
and \texttt{python-cpl} has been added for ESO pipelines and VLT data products.
For ground-based astronomy, \texttt{baseband} has added IO capabilities for
VLBI, and \texttt{SpectraPy} brings slit spectroscopy to the \astropy ecosystem.
Other new affiliated packages include \texttt{agnpy} for AGN jets,
\texttt{statmorph} for fitting galactic morphological diagnostics,
\texttt{dust\_extinction} for modeling interstellar dust extinction,
\texttt{feets} for extracting features from time series data, and
\texttt{corral} for managing data intensive parallel pipelines.
\texttt{saba} gives an interface to the Sherpa \citep{sherpa} fitting routines,
\texttt{BayesicFitting} provides an interface for generic Bayesian inference,
and \texttt{sbpy} enables calculations for asteroid and
cometary astrophysics. Finally, \texttt{synphot} provides an interface for
synthetic photometry.
% \citep{Mommert2019}

Several of the coordinated and affiliated packages described in
\cite{astropy:2018} have had substantial improvements. \texttt{astroquery}
\citep{astroquery} has added access to roughly a dozen new missions and data
services, including the James Webb Space Telescope (JWST) archive, and the
project has switched to a continuous release model: every time a change is
committed to the main development branch it is published on the Python Package
Index (PyPI) and available for installation. (However, formal releases are still
done a few times per year.) \texttt{photutils} \citep{photutils} released its
first stable version, indicating that the API will change less frequently, and
there have been several significant performance improvements. \texttt{ccdproc}
\citep{ccdproc} also released a new major version, bringing better performance to
some image combination operations. The \texttt{regions} package
\citep{pyregions}, for manipulating ds9-style region definitions \citep{ds9},
added new ways to manipulate regions and introduced new region types.
\texttt{reproject}'s \citep{reproject} major new feature is a function to align
and co-add images to create a mosaic; better support for parallelization was
also added. \texttt{specutils}, the package that defines containers for 1D and
2D spectra \citep{specutils}, also had its first stable release and the addition
of classes to read JWST data. \texttt{stingray} has also had a major performance
overhaul. The package \texttt{gammapy} \citep{gammapy} also has major performance
improvements and has unified its API in preparation for its first stable
release.

>>>>

As part of the process of developing \astropypkg into a long-term sustainable
product, and to improve transparency and accountability, the Project agreed to
write down and formalize our governance structure (partly supported by explicit
funding for this purpose; see Section~\ref{sec:project-funding}). At the 2019
Astropy Coordination Meeting, input was gathered from participants on what governance
structures existed in the associated Open Source Software communities, and what
would fit well with the needs of \astropy. This led to a ``retreat'' planned
for March 2020, but due to the COVID-19 pandemic, this became a series of
virtual meetings of the ``Astropy Governance Working Group.'' This group
drafted the APE 0 document \citep{ape0}, which was eventually ratified and
implemented by the ``Astropy Governance Implementation Working Group'' in Fall
2021. While the process emphasized flexibility and adaptability,
it is the expectation that \astropy's governance will operate in
this framework for at least the medium-term future.

The APE 0 \citep{ape0} document lays out the principles of this governance
structure, so we refer the reader to that document for a more thorough
description, and here we only highlight some key elements. The APE 0-based
governance formalizes the building blocks that are already de facto true or
previously discussed.
With this in mind, it highlights the developer and user community of the
\astropy Project as the ultimate sources of authority, as well as the core
principle of ``do-ocracy'': those who do work for the Project (coding,
training, or other harder-to-quantify contributions) gain more influence on the
outputs of the Project by virtue of their effort. However, APE 0 adds the
concept of ``voting members'': a self-governed part of that community who are
entrusted to elect the Coordination Committee (CoCo). While the CoCo
predates APE 0, membership, rights, and responsibilities are now formally outlined.
This role is mainly to facilitate consensus and make decision
when other mechanisms have failed, and includes powers that either
require central authority or secrets (e.g., passwords). However, APE 0 also
charges the Committee to devolve responsibilities and seek community input on
these items as often as possible.

The first CoCo election took place in Fall 2021, electing a mix of prior and new members.
New roles and voting members were also established. This
suggests the process is already working to serve the long-term interests of the
committee to both spread the coordination effort, and to ensure it is not
dominated by the same people for as long as the Project continues. While other,
more fine-grained governance improvements are planned for the future, it is
clear the foundation is now in place.

@@

\begin{figure}[t!]
    \begin{centering}
      \includegraphics[width=0.9\textwidth]{Ncommitters_vs_frac_commits.pdf}
        \caption{
            The cumulative fraction of Git commits to the \astropypkg core
            package as a function of number of contributors (committers) broken
            into three equal-length time periods between the start of \astropypkg
            and the feature freeze date of v5.0.
            The time eras early, mid, and recent correspond to the approximate
            year ranges 2011--2015, 2015--2018, 2018--2021 (defined more
            precisely in Section~\ref{sec:project-contributors}).
            In all eras, the majority of commits to the \astropypkg core package
            have come from a small subset of the total number of contributors.
            For example, as annotated in the figure, 90\% of the Git commits
            have come from 18, 34, and 44 committers in the early, mid, and
            recent eras, respectively.
        }
        \label{fig:contributor-summary:Ncommitters}
    \end{centering}
\end{figure}

As described in the introduction above, the \astropy Project is broader than the
core \astropypkg package as it also represents a number of core infrastructure,
coordinated, and affiliated packages, educational initiatives, and user communities.
In total, well over 1500 people have contributed in some way to the Project,
e.g., through contributing code or documentation, reporting issues
on \github, providing infrastructure support, writing tutorials, or helping to
coordinate \astropy workshops.
The vast majority of these contributors and participants have interacted as
volunteers choosing to engage with the open-source and open-development software
community.
However, while the total number of contributors is large, the number of
contributions and bulk of the maintenance of the \astropy Project components has
primarily been sustained by a small number of heavily-engaged or
explicitly-employed contributors \citep[e.g.,][]{astropy:2018}, as is the
case in many large, open-source software communities
\citep[e.g.,][]{numpy:nature}.

@@

Since its inception, the most active contributors to the \astropypkg package
have spanned a range of career stages and paths.
For example, Figure~\ref{fig:contributor-summary:top} shows a breakdown of the
career stage/path at the midpoint of the three time eras defined in the previous
paragraph for all contributors with more than 50 commits to the \astropypkg
package within each era.
Over the $\sim$6 years between the ``early'' and ``recent'' eras, many
contributors moved between these categories (i.e., have graduated as students to
become research staff/faculty or software engineers), but also a number of
early contributors moved on to other positions and stopped contributing to the
core \astropypkg package.
Given this, it is encouraging that the numbers in each career stage/path have
stayed relatively constant over the history of \astropy (with some showing
growth), as new, active contributors have generally joined and stepped in to
fill top-contributor roles within the project.
On the other hand, the total number of top contributors has not grown
substantially, reflecting one of our principal challenges: sustaining a pipeline
from users to contributors to maintainers of the package.
One other thing to note in this context is that in all eras, 50--60\% of the
software engineers who contributed to \astropypkg were (or are) employed by
STScI, whose leadership has consistently provided resources (financial and
personnel) towards sustaining \astropy.

@@

% Much of this text was adapted from Robel Geda's report from NSBP and
% SACNAS. Please include him on the author list.
\textbf{\astropy representation at national diversity conferences:}
Several members of the \astropy community attended virtual conferences of the
National Society of Black Physicists (NSBP) and SACNAS (focusing on Hispanic and
Native American scientists across all STEM fields). \astropy representatives
noted that one underlying topic came up multiple times: A major barrier for
persons who come from underrepresented communities is the lack of pedagogical
resources and expertise. For instance, software engineering skills and best
practices are not part of the standard
Physics and Astrophysics curriculum. Additionally, the larger astronomical
community is still in the process of transitioning towards \python-based tools,
hindering students not currently under advisement by someone with
extensive \python expertise from getting involved with the open-source \python
community. Projects such as Learn Astropy (see Section~\ref{sec:learn}) could
have a profound impact by empowering underrepresented groups because it provides
a free, searchable, and accessible introduction to \python tools for
astrophysical research. Representatives also noted that offering training and
teaching materials to PIs of Research Experience for Undergraduate (REU)
programs throughout the United States could be helpful for encouraging advisors
to teach students how to use Astropy, create their own libraries, and use
version control on open-source platforms like \github. Such materials could also
be offered as a workshop at national diversity conferences themselves, as SACNAS
solicits special session proposals each year.

@@

With the goal of establishing a sustainable financial model for the \astropy
Project, the CoCo, under the new governance charter (see
Section~\ref{sec:project-governance}), established a standing Finance Committee.
This committee oversees the planning and allocation of finances on behalf
of the Project. The \astropy Project accepts funds from institutions as well as
individuals through NumFOCUS.\footnote{NumFOCUS is a 501(c)(3) nonprofit that
supports and promotes world-class, innovative, open-source scientific computing.}
Previously, no direct financial support was available for project development,
and NumFOCUS covered most of the incurred operational costs. However, transitioning
to long-term financial stability meant having an influx of funds in the form of
institutional support, and hence the \astropy Project applied for and successfully
acquired four grants over the last several years. These grants are administered by NumFOCUS on behalf of the
Project, providing a ``neutral'' space not tied to any particular astronomical
institution. The two major grants are briefly summarized below.

\begin{description}
    \item[Moore] The Gordon and Betty Moore Foundation awarded the \astropy
    Project a $\sim$\$900k (US) grant in 2019 for the proposal ``Sustaining and
    Growing the Astropy Project.'' Over a three-year term, this grant supported
    sustenance of the Project by providing funds for (1) the development and
    maintenance of \astropypkg and its infrastructure, provided to existing
    project members and targeted hires, (2) the transition of long-term users to
    contributors and maintainers through mentorship efforts, (3) the
    formalization of a governance structure (see
    Section~\ref{sec:project-governance}), and (4) the improvement of
    IDE efforts in the project (see
    Section~\ref{sec:project-ide}). Some of this money supported travel for
    meetings, conferences, and workshops. This generous grant helped reduce
    reliance on volunteer-driven maintenance of the openly-developed package and
    paved the way to acquire funding from US federal agencies, representing a
    major milestone for the success of the project.

    \item[NASA] A successful proposal to the NASA ROSES-2020
    call\footnote{http://solicitation.nasaprs.com/ROSES2020}, section E.7
    (Support for Open Source Tools, Frameworks, and Libraries), granted the
    \astropy Project $\sim$\$600k (US) to support its work. Funded work for a
    term of three years includes (1) project-wide infrastructure maintenance and
    improvements, (2) targeted work on areas of the \astropy core and
    coordinated packages, (3) enhancements to the Learn
    Astropy ecosystem and educational materials (see Section~\ref{sec:learn}),
    and (4) support for the \astropy affiliated packages. These goals
    support the sustenance of the project by encouraging further
    development of the package ecosystem and community engagement.
\end{description}

In addition to the two grants described above, the Project benefited from
a Gemini Observatory contract under its Science User Support Department,
which awarded funds for development work that supports both the Astropy and
DRAGONS projects. The Project also received financial support from the Dunlap
Seed Funding program at the University of Toronto to develop educational
resources for scientific software packaging within the Learn Astropy
framework. The project also indirectly benefited from the NRAO ALMA Development
Study program, which supported the grant ``Linking CASA to the astropy
ecosystem''\footnote{\url{https://science.nrao.edu/facilities/alma/science_sustainability/Spectral_Cube_and_Radio_Beam_Ginsburg.pdf}}.
All of these successful proposals show that the \astropy Project
is vital for the astronomical infrastructure and community at large, and
strengthen its promising and sustainable future.

@@

As defined in \paperii, Astropy affiliated packages are astronomy-related
\python packages that provide functionality that builds upon and extends the
\astropypkg core package that have requested to be included as part of the
Astropy Project community. These packages support the goals and vision of
Astropy of improving code re-use, interoperability, and embracing good coding
practices such as testing and thorough documentation.
All Astropy-coordinated and other affiliated packages are listed with detailed
information on the \href{{https://www.astropy.org/affiliated}}{\astropy
website}; a table summarizing key aspects of the affiliated packages is included
in the Appendix below (Table~\ref{tab:affiliated-registry}).

Since \paperii, there has been an expansion of affiliated
packages for gravitational astrophysics, including:
\texttt{PyCBC} for exploring gravitational wave signals, \texttt{lenstronomy} for
modeling strong gravitational lenses, \texttt{ligo.skymap} for visualizing
gravitational wave probability maps, and \texttt{EinsteinPy} for general
relativity and gravitational astronomy. There have also been several packages
added to the ecosystem related to HEALPix: \texttt{astropy-healpix}
for a BSD-licensed HEALPix implementation, and \texttt{mocpy} for Multi-Order
Coverage maps. \texttt{astroalign} has been introduced for astrometric registration,
and \texttt{python-cpl} has been added for ESO pipelines and VLT data products.
For ground-based astronomy, \texttt{baseband} has added IO capabilities for
VLBI, and \texttt{SpectraPy} brings slit spectroscopy to the \astropy ecosystem.
Other new affiliated packages include \texttt{agnpy} for AGN jets,
\texttt{statmorph} for fitting galactic morphological diagnostics,
\texttt{dust\_extinction} for modeling interstellar dust extinction,
\texttt{feets} for extracting features from time series data, and
\texttt{corral} for managing data intensive parallel pipelines.
\texttt{casaformats-io} provides infrastructure for reading data stored in CASA table formats.
\texttt{saba} gives an interface to the Sherpa \citep{sherpa} fitting routines,
\texttt{BayesicFitting} provides an interface for generic Bayesian inference,
and \texttt{sbpy} enables calculations for asteroid and
cometary astrophysics. Finally, \texttt{synphot} provides an interface for
synthetic photometry.
% \citep{Mommert2019}

Several of the coordinated and affiliated packages described in
\cite{astropy:2018} have had substantial improvements. \texttt{astroquery}
\citep{astroquery} has added access to roughly a dozen new missions and data
services, including the James Webb Space Telescope (JWST) archive, and the
project has switched to a continuous release model: every time a change is
committed to the main development branch it is published on the Python Package
Index (PyPI) and available for installation. (However, formal releases are still
done a few times per year.) \texttt{photutils} \citep{photutils} released its
first stable version, indicating that the API will change less frequently, and
there have been several significant performance improvements. \texttt{ccdproc}
\citep{ccdproc}  released a new major version, bringing better performance to
some image combination operations. The \texttt{regions} package
\citep{pyregions}, for manipulating ds9-, CASA-, or FITS-style region definitions \citep{ds9},
added new ways to manipulate regions, introduced new region types, and obtained
expanded visualization and interactive region editing capabilities.
\texttt{reproject}'s \citep{reproject} major new feature is a function to align
and co-add images to create a mosaic; better support for parallelization was
also added. \texttt{specutils}, the package that defines containers for 1D and
2D spectra \citep{specutils},  had its first stable release and the addition
of classes to read JWST data. \texttt{stingray} has had a major performance
overhaul. The package \texttt{gammapy} \citep{gammapy} has had major performance
improvements and has unified its API in preparation for its first stable
release.
